---
title: "The Brazilian eCommerce"
author: Daniel Rabi and Jaswinder Brar
date: June 10, 2026
format: html
editor: visual
embed-resources: TRUE
toc: TRUE

echo: FALSE
message: FALSE
warning: FALSE
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(DBI)
library(RSQLite)
library(lubridate)
library(tsibble)
library(scales)
library(feasts)
library(fabletools)
library(patchwork)
```

```{r}
data <- DBI::dbConnect(RSQLite::SQLite(), "olist.sqlite")

customers <- DBI::dbReadTable(data, "customersT")
geo_locations <- DBI::dbReadTable(data, "geolocationsT")
order_items <- DBI::dbReadTable(data, "order_itemsT")
order_payments <- DBI::dbReadTable(data, "order_paymentsT")
order_reviews <- DBI::dbReadTable(data, "order_reviewsT")
orders <- DBI::dbReadTable(data, "ordersT")
product_translations <- DBI::dbReadTable(data, "product_category_name_translationsT")
products <- DBI::dbReadTable(data, "productsT")
sellers <- DBI::dbReadTable(data, "sellersT")

DBI::dbDisconnect(data)
```

```{r}
# Cleaning missing values
clean_missing <- function(data){
  data_clean <-  data %>% 
    dplyr::mutate(dplyr::across(tidyselect::where(is.character), ~ dplyr::na_if(stringr::str_trim(.x), "")))
  
  return(data_clean)
}
```

```{r, warning=FALSE}
# Cleaning the data
customers <- customers %>% 
  dplyr::mutate(
    dplyr::across(3, as.numeric)
  ) %>% 
  clean_missing()

geo_locations <- geo_locations %>% 
  dplyr::mutate(
    dplyr::across(1:3, as.numeric)
  ) %>% 
  clean_missing()

order_items <- order_items %>% 
  dplyr::mutate(
    dplyr::across(c(order_item_id, price, freight_value), as.numeric),
    dplyr::across(shipping_limit_date, lubridate::ymd_hms)
  ) %>% 
  clean_missing()

order_payments <- order_payments %>% 
  dplyr::mutate(
    dplyr::across(c(2, 4, 5), as.numeric)
  ) %>% 
  clean_missing()

order_reviews <- order_reviews %>% 
  dplyr::mutate(
    dplyr::across(3, as.numeric),
    dplyr::across(6:7, lubridate::ymd_hms)
  ) %>% 
  clean_missing()

orders <- orders %>% 
  dplyr::mutate(
    dplyr::across(4:8, lubridate::ymd_hms)
  )

product_translations <- product_translations %>% clean_missing()

products <- products %>% 
  dplyr::mutate(
    dplyr::across(3:9, as.numeric)
  ) %>% 
  clean_missing()

sellers <- sellers %>% 
  dplyr::mutate(
    dplyr::across(2, as.numeric)
  ) %>% 
  clean_missing()
```

```{r, include=FALSE}
customers
geo_locations
order_items
order_payments
order_reviews
orders
product_translations
products
sellers
```

# Things to look at:

- Seasonality of purchases - is there seasonality?

  - What day of the week sales the most

  - What are the best hours of the day for sales

- Location of purchases

- Most common type of payment

- Who is the best seller right now

- Who are we expecting to be the best seller in the future

  - Forecasting reviews ratings and sales for each seller to tell us who we think will be the best seller

  - I think we can use both forecasts to help us understand who's the best seller rather than making a decision on which one of the two defines what best is

- What is the best selling product

  - Forecast which product we expect to sell the best

- How many returning customers does each seller has? Does it correlate with their reviews

# Seasonality

```{r}
combined_purchase <- orders %>% 
  dplyr::transmute(
    order_id,
    customer_id,
    purchase_datetime = order_purchase_timestamp,
    purchase_date = as.Date(purchase_datetime),
    purchase_month = lubridate::month(purchase_datetime, label = TRUE),
    purchase_day = lubridate::wday(purchase_datetime, label = TRUE),
    purchase_yearmonth = tsibble::yearmonth(purchase_date)
    ) %>% 
  dplyr::left_join(
    order_payments,
    by = "order_id"
  )

seasonality_purchase <- combined_purchase %>% 
  dplyr::group_by(purchase_yearmonth) %>% 
  dplyr::summarise(
    total_value = sum(payment_value, na.rm = TRUE),
    total_orders = dplyr::n_distinct(order_id),
    total_customers = dplyr::n_distinct(customer_id),
    first_date = min(purchase_date, na.rm = TRUE),
    last_date = max(purchase_date, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  tsibble::as_tsibble(index = purchase_yearmonth)
```

```{r, fig.height= 9, fig.width= 8}
seas_1 <- seasonality_purchase %>% 
  tsibble::fill_gaps(total_value = 0) %>% 
  ggplot2::autoplot(total_value) +
  ggplot2::scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  ggplot2::labs(
    title = "Monthly Total Purchase Value Over Time",
    x = "Time",
    y = "Total Value ($)"
  ) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))

seas_2 <- seasonality_purchase %>% 
  tsibble::fill_gaps(total_value = 0) %>% 
  feasts::gg_subseries(total_value) +
  tsibble::scale_x_yearmonth(date_labels = "%Y") +
  ggplot2::labs(
    title = "Total Value of Purchases Over Time",
    subtitle = "Highliting the trend by Year",
    x = "Time",
    y = "Total Value ($)"
  ) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5)) +
  ggplot2::theme(plot.subtitle = ggplot2::element_text(hjust = 0.5))

stl_model <- seasonality_purchase %>% 
  tsibble::fill_gaps(total_value = 0) %>% 
  fabletools::model(feasts::STL(formula = total_value ~ season(window = 13)))

seas_3 <- stl_model %>% 
  fabletools::components() %>% 
  ggplot2::autoplot() +
  ggplot2::scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  ggplot2::labs(
    title = "STL Decomposition of Monthly Purchase Value",
    subtitle = "Total Value = Trend + Season Year + Remainder",
    x = "Time",
    y = "Total Value ($)"
  ) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5)) +
  ggplot2::theme(plot.subtitle = ggplot2::element_text(hjust = 0.5))

seas_1 / seas_3
```

The graph above displays an upward trend in total value of purchases and rejects the idea of seasonality across total value of purchases. The suddent drop in total value in July 2018 arises from the simple reason that the data we possess ends at July 5th, 2018 and not from a drop in purchases.
