---
title: "The Brazilian eCommerce"
author: Daniel Rabi and Jaswinder Brar
date: June 10, 2026
format: html
editor: visual
embed-resources: TRUE
toc: TRUE

echo: FALSE
message: FALSE
warning: FALSE
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(DBI)
library(RSQLite)
library(tsibble)
library(scales)
library(feasts)
library(fabletools)
library(patchwork)
library(slider)
library(tidymodels)
library(rsample)
```

```{r}
data <- DBI::dbConnect(RSQLite::SQLite(), "olist.sqlite")

customers <- DBI::dbReadTable(data, "customersT")
geo_locations <- DBI::dbReadTable(data, "geolocationsT")
order_items <- DBI::dbReadTable(data, "order_itemsT")
order_payments <- DBI::dbReadTable(data, "order_paymentsT")
order_reviews <- DBI::dbReadTable(data, "order_reviewsT")
orders <- DBI::dbReadTable(data, "ordersT")
product_translations <- DBI::dbReadTable(data, "product_category_name_translationsT")
products <- DBI::dbReadTable(data, "productsT")
sellers <- DBI::dbReadTable(data, "sellersT")

DBI::dbDisconnect(data)
```

```{r}
# Cleaning missing values
clean_missing <- function(data){
  data_clean <-  data %>% 
    dplyr::mutate(dplyr::across(tidyselect::where(is.character), ~ dplyr::na_if(stringr::str_trim(.x), "")))
  
  return(data_clean)
}
```

```{r, warning=FALSE}
# Cleaning the data
customers <- customers %>% 
  dplyr::mutate(
    dplyr::across(3, as.numeric)
  ) %>% 
  clean_missing()

geo_locations <- geo_locations %>% 
  dplyr::mutate(
    dplyr::across(1:3, as.numeric)
  ) %>% 
  clean_missing()

order_items <- order_items %>% 
  dplyr::mutate(
    dplyr::across(c(order_item_id, price, freight_value), as.numeric),
    dplyr::across(shipping_limit_date, lubridate::ymd_hms)
  ) %>% 
  clean_missing()

order_payments <- order_payments %>% 
  dplyr::mutate(
    dplyr::across(c(2, 4, 5), as.numeric)
  ) %>% 
  clean_missing()

order_reviews <- order_reviews %>% 
  dplyr::mutate(
    dplyr::across(3, as.numeric),
    dplyr::across(6:7, lubridate::ymd_hms)
  ) %>% 
  clean_missing()

orders <- orders %>% 
  dplyr::mutate(
    dplyr::across(4:8, lubridate::ymd_hms),
    late = dplyr::case_when(
      order_estimated_delivery_date >= order_delivered_customer_date ~ 0, # 0: FALSE
      order_estimated_delivery_date < order_delivered_customer_date ~ 1 # 1: TRUE
    )
  )

product_translations <- product_translations %>% clean_missing()

products <- products %>% 
  dplyr::mutate(
    dplyr::across(3:9, as.numeric)
  ) %>% 
  clean_missing()

sellers <- sellers %>% 
  dplyr::mutate(
    dplyr::across(2, as.numeric)
  ) %>% 
  clean_missing()
```

```{r, include=FALSE}
customers
geo_locations
order_items
order_payments
order_reviews
orders
product_translations
products
sellers
```

# Things to look at:

- Seasonality of purchases - is there seasonality?

  - What day of the week sales the most

  - What are the best hours of the day for sales

- Location of purchases

- Most common type of payment

- Who is the best seller right now

- Who are we expecting to be the best seller in the future

  - Predicting reviews ratings and sales for each seller to tell us who we think will be the best seller

  - I think we can use both forecasts to help us understand who's the best seller rather than making a decision on which one of the two defines what best is

- What is the best selling product category

  - Predict which category we expect to sell the best next month

  - Factors:

    - Lag revenue

    - Review scores

    - Review comment no comment

    - Rolling SD

    - Rolling mean

    - 

- How many returning customers does each seller has? Does it correlate with their reviews

# Seasonality

```{r}
combined_purchase <- orders %>% 
  dplyr::transmute(
    order_id,
    customer_id,
    purchase_datetime = order_purchase_timestamp,
    purchase_date = as.Date(purchase_datetime),
    purchase_month = lubridate::month(purchase_datetime, label = TRUE),
    purchase_day = lubridate::wday(purchase_datetime, label = TRUE),
    purchase_yearmonth = tsibble::yearmonth(purchase_date),
    purchase_yearweek = tsibble::yearweek(purchase_date)
    ) %>% 
  dplyr::left_join(
    order_payments %>% 
      dplyr::group_by(order_id),
    by = "order_id"
  )

seasonality_purchase <- combined_purchase %>% 
  dplyr::group_by(purchase_yearweek) %>% 
  dplyr::summarise(
    total_value = sum(payment_value, na.rm = TRUE),
    total_orders = dplyr::n_distinct(order_id),
    total_customers = dplyr::n_distinct(customer_id),
    first_date = min(purchase_date, na.rm = TRUE),
    last_date = max(purchase_date, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  tsibble::as_tsibble(index = purchase_yearweek)
```

```{r}
seas_1 <- seasonality_purchase %>% 
  tsibble::fill_gaps(total_value = 0) %>% 
  ggplot2::autoplot(total_value) +
  ggplot2::scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  ggplot2::labs(
    title = "Monthly Total Purchase Value Over Time",
    x = "Time",
    y = "Total Value ($)"
  ) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))

seas_2 <- seasonality_purchase %>% 
  tsibble::fill_gaps(total_value = 0) %>% 
  feasts::gg_subseries(total_value) +
  tsibble::scale_x_yearweek(date_labels = "%Y") +
  ggplot2::labs(
    title = "Total Value of Purchases Over Time",
    subtitle = "Highliting the trend by Year",
    x = "Time",
    y = "Total Value ($)"
  ) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5)) +
  ggplot2::theme(plot.subtitle = ggplot2::element_text(hjust = 0.5))

stl_model <- seasonality_purchase %>% 
  tsibble::fill_gaps(total_value = 0) %>% 
  fabletools::model(feasts::STL(formula = total_value ~ season(window = 13)))

seas_3 <- stl_model %>% 
  fabletools::components() %>% 
  ggplot2::autoplot() +
  ggplot2::scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  tsibble::scale_x_yearweek(
    date_breaks = "3 months",
    date_labels = "%Y %b"
    ) +
  ggplot2::labs(
    title = "STL Decomposition of Weekly Purchase Value",
    subtitle = "Total Value = Trend + Season Year + Remainder",
    x = "Time",
    y = "Total Value ($)"
  ) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5)) +
  ggplot2::theme(plot.subtitle = ggplot2::element_text(hjust = 0.5))

seas_3
```

The graph above displays an upward trend in total value of purchases and rejects the idea of seasonality across total value of purchases. The sudden drop in total value in July 2018 arises from the simple reason that the data we possess ends at July 5th, 2018 and not from a drop in purchases.

# Best Seller

```{r}
combined_purchase
seasonality_purchase
```

```{r}
sellers_tbl <- sellers %>% 
  dplyr::left_join(
    order_items,
    by = "seller_id"
  ) %>% 
  dplyr::group_by(seller_id) %>% 
  dplyr::summarise(total_price = sum(price))

sellers_tbl
```

```{r}
sellers_tbl %>% 
  dplyr::arrange(desc(total_price)) %>% 
  dplyr::slice_head(n = 5) %>% 
  ggplot2::ggplot(ggplot2::aes(x = seller_id, y = total_price)) +
  ggplot2::geom_col(ggplot2::aes(fill = seller_id))
```

# Best Product Category

```{r}
products %>% 
  dplyr::filter(is.na(product_category_name) == TRUE)
```

```{r}
product_tbl <- products %>% 
  dplyr::left_join(
    order_items,
    by = "product_id"
  ) %>% 
  dplyr::left_join(
    product_translations,
    by = "product_category_name"
  ) %>% 
  dplyr::group_by(product_category_name_english) %>% 
  dplyr::select(-product_category_name) %>% 
  dplyr::rename(product_category = product_category_name_english)

product_tbl
```

```{r}
product_tbl %>% 
  dplyr::filter(is.na(product_category) == TRUE)
```

```{r}
best_category <- product_tbl %>% 
  dplyr::summarise(
    total_revenue = sum(price, na.rm = TRUE),
    total_sold = n(),
    .groups = "drop"
  ) %>% 
  dplyr::arrange(desc(total_revenue))

best_category %>% 
  dplyr::slice_head(n = 5) %>% 
  ggplot2::ggplot(ggplot2::aes(x = stats::reorder(product_category, -total_revenue), y = total_revenue, fill = product_category)) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -25, hjust = -0.01)) +
  ggplot2::labs(
    title = "Top 5 Total Revenue by Category",
    x = "Product Category",
    y = "Total Revenue",
    fill = "Product Category"
  ) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
```

As we can see above, the top five best selling product categories are:

1.  Health beauty
2.  Watches gifts
3.  Bed bath tables
4.  Sports leisure
5.  Computers accessories

However, the fact that the health beauty category is best seller does not guarantee that it will stay the best seller. Therefore we will use machine learning to predict which category we expect to be the best selling product category next month.

## Predicting Weekly Category Revenue

Eight orders were identified as "delivered" but lacked a recorded customer delivery date. Since delivery timeliness could not be determined for these observations, they were excluded from the late-delivery analysis. Missing weekly review scores were imputed using the average review score within the corresponding product category. For category-weeks with no observed delivery records, the late-delivery proportion was set to zero.

```{r}
weekly_reviews <- order_reviews %>%
  dplyr::left_join(
    orders %>% 
      dplyr::select(order_id, order_purchase_timestamp),
    by = "order_id"
  ) %>% 
  dplyr::group_by(order_id) %>%
  dplyr::summarise(
    purchase_yearweek = tsibble::yearweek(max(order_purchase_timestamp, na.rm = TRUE)),
    review_score = mean(review_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::left_join(
    product_tbl %>%
      dplyr::select(order_id, product_category) %>%
      dplyr::distinct() %>%
      tidyr::drop_na(product_category),
    by = "order_id"
  ) %>%
  tidyr::drop_na(product_category) %>% 
  dplyr::group_by(product_category) %>%
  tidyr::complete(
    purchase_yearweek = seq(min(purchase_yearweek), max(purchase_yearweek), by = 1),
    fill = list(review_score = NA)
  ) %>%
  dplyr::mutate(
    review_score = tidyr::replace_na(review_score, mean(review_score, na.rm = TRUE))
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(purchase_yearweek, product_category) %>% 
  dplyr::summarise(
    avg_reviews = mean(review_score, na.rm = TRUE),
    .groups = "drop"
  )

weekly_orders <- orders %>% 
  dplyr::select(order_id, order_purchase_timestamp) %>% 
  dplyr::mutate(
    purchase_yearweek = tsibble::yearweek(order_purchase_timestamp)
  )

weekly_late <- orders %>%
  dplyr::filter(order_status == "delivered" & ! is.na(late)) %>% 
  dplyr::select(order_id, order_purchase_timestamp, late) %>% 
  dplyr::mutate(
    purchase_yearweek = tsibble::yearweek(order_purchase_timestamp)
  ) %>% 
  dplyr::left_join(
    product_tbl %>% 
      dplyr::select(order_id, product_category),
    by = "order_id"
  ) %>% 
  tidyr::drop_na(product_category) %>% 
  dplyr::group_by(product_category) %>% 
  dplyr::arrange(purchase_yearweek, product_category) %>%
  tidyr::complete(
    purchase_yearweek = seq(min(purchase_yearweek), max(purchase_yearweek), by = 1),
    fill = list(late = NA)
  ) %>% 
  dplyr::mutate(
    late = tidyr::replace_na(late, 0)
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(purchase_yearweek, product_category) %>% 
  dplyr::summarise(
    late_proportion = mean(late, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
pred_category <- product_tbl %>% 
  dplyr::left_join(
    weekly_orders,
    by = "order_id"
  ) %>% 
  tidyr::drop_na(product_category, purchase_yearweek) %>% 
  dplyr::group_by(purchase_yearweek, product_category) %>% 
  dplyr::summarise(
    total_revenue = sum(price, na.rm = TRUE),
    total_sellers = n_distinct(seller_id),
    .groups = "drop"
  ) %>% 
  dplyr::group_by(product_category) %>% 
  tidyr::complete(
    purchase_yearweek = seq(min(purchase_yearweek), max(purchase_yearweek), by = 1),
    fill = list(
      total_revenue = 0,
      total_sellers = 0
    )
  ) %>% 
  dplyr::left_join(
    weekly_reviews,
    by = c("purchase_yearweek", "product_category")
  ) %>% 
  dplyr::left_join(
    weekly_late,
    by = c("purchase_yearweek", "product_category")
  ) %>% 
  dplyr::mutate(late_proportion = tidyr::replace_na(late_proportion, 0)) %>% 
  dplyr::arrange(product_category, purchase_yearweek) %>% 
  dplyr::group_by(product_category) %>% 
  dplyr::mutate(
    lag_revenue = dplyr::lag(total_revenue, 1),
    rolling_sd4 = slider::slide_dbl(
      .x = total_revenue,
      .f = sd,
      .before = 4, # Looking 4 weeks back
      .after = 0,
      .complete = FALSE 
    ),
    rolling_mean4 = slider::slide_dbl(
      .x = total_revenue,
      .f = mean,
      .before = 4,
      .after = 0,
      .complete = FALSE 
    )
  ) %>% 
  tidyr::drop_na(lag_revenue) %>% 
  dplyr::ungroup()

pred_category
```

```{r}
pred_category_tbl <- pred_category %>% 
  dplyr::arrange(product_category, purchase_yearweek) %>% 
  dplyr::group_by(product_category) %>% 
  dplyr::mutate(
    pred_date = dplyr::lead(purchase_yearweek, n = 1),
    future_revenue = dplyr::lead(total_revenue, n = 1) # Adding the future revenue to each row so that the model knows what the actual result was so that he can learn the behaviour
  ) %>% 
  tidyr::drop_na(pred_date) %>% 
  dplyr::ungroup()

pred_category_tbl
```

```{r}
# Sampling
split_data <- pred_category_tbl %>% 
  rsample::initial_time_split(prop = 0.7)

training_data <- rsample::training(split_data)
testing_data <- rsample::testing(split_data)

# Features Engineering
recipe_pipeline <- 
  recipes::recipe(future_revenue ~ ., data = training_data) %>% 
  recipes::step_rm(purchase_yearweek, pred_date, product_category) %>% 
  recipes::step_normalize(all_numeric_predictors()) %>% 
  recipes::prep()

training_baked <- recipes::bake(recipe_pipeline, training_data)
testing_baked <- recipes::bake(recipe_pipeline, testing_data)


# Modeling - Linear Regression
model_lm <-
  parsnip::linear_reg(mode = "regression") %>% 
  parsnip::set_engine("lm") %>% 
  parsnip::fit(future_revenue ~ ., data = training_baked)

# Predicting
predicted_lm <- model_lm %>% 
  stats::predict(new_data = testing_baked) %>% 
  dplyr::bind_cols(testing_data %>% dplyr::select(product_category, purchase_yearweek, future_revenue))

# Metrics
tidy_lm <- parsnip::tidy(model_lm) %>% 
  gt::gt() %>% 
  gt::fmt_percent(columns = matches("std.error"), decimals = 2)

glance_lm <- parsnip::glance(model_lm) %>% 
  gt::gt() %>% 
  gt::fmt_percent(columns = matches("r.squared|adj.r.squared"), decimals = 2)

yardstick_lm <- predicted_lm %>% 
  yardstick::metrics(truth = future_revenue, estimate = .pred) %>% 
  dplyr::arrange(.metric) %>% 
  gt::gt() %>% 
  gt::fmt_percent(columns = matches(".estimate"), decimals = 2)
```

```{r}
yardstick_lm
glance_lm
```

```{r}
multi_regression <- function(data_table){
  dependent <- names(data_table)[2] # This is what we're trying to predict. The y
  independents <- names(data_table)[-(1:2)] # The predictors
  
  regression_models <- dplyr::tibble(model = "") # Created an empty table so I can add to it later
  
  for (i in 1:length(independents)){
    betas <- paste(independents[1:i], collapse = "+")
    regression_models[i, 1] <- glue::glue("{dependent} ~ {betas}")
  }
  
  regression_models <- regression_models %>% 
    dplyr::mutate(
      regression = purrr::map(model, ~stats::lm(stats::as.formula(.x), data = data_table)),
      statistics = purrr::map(regression, broom::glance)
    ) %>% 
    tidyr::unnest(statistics) %>% 
    dplyr::select(1:4, p.value) %>% 
    dplyr::mutate_if(is.numeric, round, 3)
  
  return(regression_models)
}

multi_regression(return_wide)
```

## Predict Weekly Category Order Volume

```{r}
pred_orders <- product_tbl %>% 
  dplyr::left_join(
    weekly_orders,
    by = "order_id"
  ) %>% 
  dplyr::group_by(purchase_yearweek, product_category) %>% 
  dplyr::summarise(
    total_orders = dplyr::n_distinct(order_id),
    total_revenue = sum(price, na.rm = TRUE),
    avg_order_value = total_revenue / total_orders,
    total_sellers = dplyr::n_distinct(seller_id),
    .groups = "drop"
  ) %>% 
  dplyr::arrange(product_category, purchase_yearweek) %>% 
  dplyr::group_by(product_category) %>% 
  dplyr::mutate(
    lag_orders = dplyr::lag(total_orders, 1),
    rolling_mean4 = slider::slide_dbl(
      total_orders,
      mean,
      .before = 4,
      .after = 0,
      .complete = FALSE
    ),
    rolling_sd4 = slider::slide_dbl(
      total_orders,
      sd,
      .before = 4,
      .after = 0,
      .complete = FALSE
    )
  ) %>% 
  tidyr::drop_na() %>% 
  dplyr::ungroup()

pred_orders
```

# Weekly Customer Retention Rate

What are the chances that a customer will purchase again

Predictors:

- Lag 1 for whether a customer bought or not

- Left a comment or not

- Average review score

- Was delivery delayed?

Target:

- 
