---
title: "The Brazilian eCommerce"
author: Daniel Rabi and Jaswinder Brar
date: June 10, 2026
format: html
editor: visual
embed-resources: TRUE
toc: TRUE

echo: FALSE
message: FALSE
warning: FALSE
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(DBI)
library(RSQLite)
library(tsibble)
library(scales)
library(feasts)
library(fabletools)
library(patchwork)
library(slider)
library(tidymodels)
library(rsample)
library(rpart.plot)
library(gt)
```

```{r}
options(mc.cores = parallel::detectCores())
```

```{r}
data <- DBI::dbConnect(RSQLite::SQLite(), "olist.sqlite")

customers <- DBI::dbReadTable(data, "customersT")
geo_locations <- DBI::dbReadTable(data, "geolocationsT")
order_items <- DBI::dbReadTable(data, "order_itemsT")
order_payments <- DBI::dbReadTable(data, "order_paymentsT")
order_reviews <- DBI::dbReadTable(data, "order_reviewsT")
orders <- DBI::dbReadTable(data, "ordersT")
product_translations <- DBI::dbReadTable(data, "product_category_name_translationsT")
products <- DBI::dbReadTable(data, "productsT")
sellers <- DBI::dbReadTable(data, "sellersT")

DBI::dbDisconnect(data)
```

```{r}
# Cleaning missing values
clean_missing <- function(data){
  data_clean <-  data %>% 
    dplyr::mutate(dplyr::across(tidyselect::where(is.character), ~ dplyr::na_if(stringr::str_trim(.x), "")))
  
  return(data_clean)
}
```

```{r, warning=FALSE}
# Cleaning the data
customers <- customers %>% 
  dplyr::mutate(
    dplyr::across(3, as.numeric)
  ) %>% 
  clean_missing()

geo_locations <- geo_locations %>% 
  dplyr::mutate(
    dplyr::across(1:3, as.numeric)
  ) %>% 
  clean_missing()

order_items <- order_items %>% 
  dplyr::mutate(
    dplyr::across(c(order_item_id, price, freight_value), as.numeric),
    dplyr::across(shipping_limit_date, lubridate::ymd_hms)
  ) %>% 
  clean_missing()

order_payments <- order_payments %>% 
  dplyr::mutate(
    dplyr::across(c(2, 4, 5), as.numeric)
  ) %>% 
  clean_missing()

order_reviews <- order_reviews %>% 
  dplyr::mutate(
    dplyr::across(3, as.numeric),
    dplyr::across(6:7, lubridate::ymd_hms)
  ) %>% 
  clean_missing()

orders <- orders %>% 
  dplyr::mutate(
    dplyr::across(4:8, lubridate::ymd_hms),
    late = dplyr::case_when(
      order_estimated_delivery_date >= order_delivered_customer_date ~ 0, # 0: FALSE
      order_estimated_delivery_date < order_delivered_customer_date ~ 1 # 1: TRUE
    )
  )

product_translations <- product_translations %>% clean_missing()

products <- products %>% 
  dplyr::mutate(
    dplyr::across(3:9, as.numeric)
  ) %>% 
  clean_missing()

sellers <- sellers %>% 
  dplyr::mutate(
    dplyr::across(2, as.numeric)
  ) %>% 
  clean_missing()
```

```{r, include=FALSE}
customers
geo_locations
order_items
order_payments
order_reviews
orders
product_translations
products
sellers
```

# Things to look at:

- Seasonality of purchases - is there seasonality?

  - What day of the week sales the most

  - What are the best hours of the day for sales

- Location of purchases

- Most common type of payment

- Who is the best seller right now

- Who are we expecting to be the best seller in the future

  - Predicting reviews ratings and sales for each seller to tell us who we think will be the best seller

  - I think we can use both forecasts to help us understand who's the best seller rather than making a decision on which one of the two defines what best is

- What is the best selling product category

  - Predict which category we expect to sell the best next month

  - Factors:

    - Lag revenue

    - Review scores

    - Review comment no comment

    - Rolling SD

    - Rolling mean

    - 

- How many returning customers does each seller has? Does it correlate with their reviews

# Seasonality

```{r}
combined_purchase <- orders %>% 
  dplyr::transmute(
    order_id,
    customer_id,
    purchase_datetime = order_purchase_timestamp,
    purchase_date = as.Date(purchase_datetime),
    purchase_month = lubridate::month(purchase_datetime, label = TRUE),
    purchase_day = lubridate::wday(purchase_datetime, label = TRUE),
    purchase_yearmonth = tsibble::yearmonth(purchase_date),
    purchase_yearweek = tsibble::yearweek(purchase_date)
    ) %>% 
  dplyr::left_join(
    order_payments %>% 
      dplyr::group_by(order_id),
    by = "order_id"
  )

seasonality_purchase <- combined_purchase %>% 
  dplyr::group_by(purchase_yearweek) %>% 
  dplyr::summarise(
    total_value = sum(payment_value, na.rm = TRUE),
    total_orders = dplyr::n_distinct(order_id),
    total_customers = dplyr::n_distinct(customer_id),
    first_date = min(purchase_date, na.rm = TRUE),
    last_date = max(purchase_date, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  tsibble::as_tsibble(index = purchase_yearweek)
```

```{r}
seas_1 <- seasonality_purchase %>% 
  tsibble::fill_gaps(total_value = 0) %>% 
  ggplot2::autoplot(total_value) +
  ggplot2::scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  ggplot2::labs(
    title = "Monthly Total Purchase Value Over Time",
    x = "Time",
    y = "Total Value ($)"
  ) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))

seas_2 <- seasonality_purchase %>% 
  tsibble::fill_gaps(total_value = 0) %>% 
  feasts::gg_subseries(total_value) +
  tsibble::scale_x_yearweek(date_labels = "%Y") +
  ggplot2::labs(
    title = "Total Value of Purchases Over Time",
    subtitle = "Highliting the trend by Year",
    x = "Time",
    y = "Total Value ($)"
  ) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5)) +
  ggplot2::theme(plot.subtitle = ggplot2::element_text(hjust = 0.5))

stl_model <- seasonality_purchase %>% 
  tsibble::fill_gaps(total_value = 0) %>% 
  fabletools::model(feasts::STL(formula = total_value ~ season(window = 13)))

seas_3 <- stl_model %>% 
  fabletools::components() %>% 
  ggplot2::autoplot() +
  ggplot2::scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  tsibble::scale_x_yearweek(
    date_breaks = "3 months",
    date_labels = "%Y %b"
    ) +
  ggplot2::labs(
    title = "STL Decomposition of Weekly Purchase Value",
    subtitle = "Total Value = Trend + Season Year + Remainder",
    x = "Time",
    y = "Total Value ($)"
  ) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5)) +
  ggplot2::theme(plot.subtitle = ggplot2::element_text(hjust = 0.5))

seas_3
```

The graph above displays an upward trend in total value of purchases and rejects the idea of seasonality across total value of purchases. The sudden drop in total value in July 2018 arises from the simple reason that the data we possess ends at July 5th, 2018 and not from a drop in purchases.

# Best Seller

```{r}
combined_purchase
seasonality_purchase
```

```{r}
sellers_tbl <- sellers %>% 
  dplyr::left_join(
    order_items,
    by = "seller_id"
  ) %>% 
  dplyr::group_by(seller_id) %>% 
  dplyr::summarise(total_price = sum(price))

sellers_tbl
```

```{r}
sellers_tbl %>% 
  dplyr::arrange(desc(total_price)) %>% 
  dplyr::slice_head(n = 5) %>% 
  ggplot2::ggplot(ggplot2::aes(x = seller_id, y = total_price)) +
  ggplot2::geom_col(ggplot2::aes(fill = seller_id))
```

# Best Product Category

```{r}
products %>% 
  dplyr::filter(is.na(product_category_name) == TRUE)
```

```{r}
product_tbl <- products %>% 
  dplyr::left_join(
    order_items,
    by = "product_id"
  ) %>% 
  dplyr::left_join(
    product_translations,
    by = "product_category_name"
  ) %>% 
  dplyr::group_by(product_category_name_english) %>% 
  dplyr::select(-product_category_name) %>% 
  dplyr::rename(product_category = product_category_name_english)
```

```{r}
product_tbl
```

```{r}
product_tbl %>% 
  dplyr::filter(is.na(product_category) == TRUE)
```

```{r}
best_category <- product_tbl %>% 
  dplyr::summarise(
    total_revenue = sum(price, na.rm = TRUE),
    total_sold = n(),
    .groups = "drop"
  ) %>% 
  dplyr::arrange(desc(total_revenue))

best_category %>% 
  dplyr::slice_head(n = 5) %>% 
  ggplot2::ggplot(ggplot2::aes(x = stats::reorder(product_category, -total_revenue), y = total_revenue, fill = product_category)) +
  ggplot2::geom_col() +
  ggplot2::scale_y_continuous(labels = scales::label_number(scale_cut = scales::cut_short_scale())) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = -25, hjust = -0.01)) +
  ggplot2::labs(
    title = "Top 5 Total Revenue by Category",
    x = "Product Category",
    y = "Total Revenue",
    fill = "Product Category"
  ) +
  ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
```

As we can see above, the top five best selling product categories are:

1.  Health beauty
2.  Watches gifts
3.  Bed bath tables
4.  Sports leisure
5.  Computers accessories

However, the fact that the health beauty category is best seller does not guarantee that it will stay the best seller. Therefore we will use machine learning to predict which category we expect to be the best selling product category next month.

## Predicting Weekly Category Revenue

Eight orders were identified as "delivered" but lacked a recorded customer delivery date. Since delivery timeliness could not be determined for these observations, they were excluded from the late-delivery analysis. Missing weekly review scores were imputed using the average review score within the corresponding product category. For category-weeks with no observed delivery records, the late-delivery proportion was set to zero.

```{r}
weekly_reviews <- order_reviews %>%
  dplyr::left_join(
    orders %>% 
      dplyr::select(order_id, order_purchase_timestamp),
    by = "order_id"
  ) %>% 
  dplyr::group_by(order_id) %>%
  dplyr::summarise(
    purchase_yearweek = tsibble::yearweek(max(order_purchase_timestamp, na.rm = TRUE)),
    review_score = mean(review_score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::left_join(
    product_tbl %>%
      dplyr::select(order_id, product_category) %>%
      dplyr::distinct() %>%
      tidyr::drop_na(product_category),
    by = "order_id"
  ) %>%
  tidyr::drop_na(product_category) %>% 
  dplyr::group_by(product_category) %>%
  tidyr::complete(
    purchase_yearweek = seq(min(purchase_yearweek), max(purchase_yearweek), by = 1),
    fill = list(review_score = NA)
  ) %>%
  dplyr::mutate(
    review_score = tidyr::replace_na(review_score, mean(review_score, na.rm = TRUE))
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(purchase_yearweek, product_category) %>% 
  dplyr::summarise(
    avg_reviews = mean(review_score, na.rm = TRUE),
    .groups = "drop"
  )

weekly_orders <- orders %>% 
  dplyr::select(order_id, order_purchase_timestamp) %>% 
  dplyr::mutate(
    purchase_yearweek = tsibble::yearweek(order_purchase_timestamp)
  )

weekly_late <- orders %>%
  dplyr::filter(order_status == "delivered" & ! is.na(late)) %>% 
  dplyr::select(order_id, order_purchase_timestamp, late) %>% 
  dplyr::mutate(
    purchase_yearweek = tsibble::yearweek(order_purchase_timestamp)
  ) %>% 
  dplyr::left_join(
    product_tbl %>% 
      dplyr::select(order_id, product_category),
    by = "order_id"
  ) %>% 
  tidyr::drop_na(product_category) %>% 
  dplyr::group_by(product_category) %>% 
  dplyr::arrange(purchase_yearweek, product_category) %>%
  tidyr::complete(
    purchase_yearweek = seq(min(purchase_yearweek), max(purchase_yearweek), by = 1),
    fill = list(late = NA)
  ) %>% 
  dplyr::mutate(
    late = tidyr::replace_na(late, 0)
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(purchase_yearweek, product_category) %>% 
  dplyr::summarise(
    late_proportion = mean(late, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
pred_category <- product_tbl %>% 
  dplyr::left_join(
    weekly_orders,
    by = "order_id"
  ) %>% 
  tidyr::drop_na(product_category, purchase_yearweek) %>% 
  dplyr::group_by(purchase_yearweek, product_category) %>% 
  dplyr::summarise(
    total_revenue = sum(price, na.rm = TRUE),
    total_sellers = n_distinct(seller_id),
    .groups = "drop"
  ) %>% 
  dplyr::group_by(product_category) %>% 
  tidyr::complete(
    purchase_yearweek = seq(min(purchase_yearweek), max(purchase_yearweek), by = 1),
    fill = list(
      total_revenue = 0,
      total_sellers = 0
    )
  ) %>% 
  dplyr::left_join(
    weekly_reviews,
    by = c("purchase_yearweek", "product_category")
  ) %>% 
  dplyr::left_join(
    weekly_late,
    by = c("purchase_yearweek", "product_category")
  ) %>% 
  dplyr::mutate(late_proportion = tidyr::replace_na(late_proportion, 0)) %>% 
  dplyr::arrange(product_category, purchase_yearweek) %>% 
  dplyr::group_by(product_category) %>% 
  dplyr::mutate(
    lag_revenue = dplyr::lag(total_revenue, 1),
    rolling_sd4 = slider::slide_dbl(
      .x = total_revenue,
      .f = sd,
      .before = 4, # Looking 4 weeks back
      .after = 0,
      .complete = FALSE 
    ),
    rolling_mean4 = slider::slide_dbl(
      .x = total_revenue,
      .f = mean,
      .before = 4,
      .after = 0,
      .complete = FALSE 
    )
  ) %>% 
  tidyr::drop_na(lag_revenue) %>% 
  dplyr::ungroup()
```

```{r}
pred_category
```

```{r}
pred_category_tbl <- pred_category %>% 
  dplyr::arrange(product_category, purchase_yearweek) %>% 
  dplyr::group_by(product_category) %>% 
  dplyr::mutate(
    pred_date = dplyr::lead(purchase_yearweek, n = 1),
    future_revenue = dplyr::lead(total_revenue, n = 1) # Adding the future revenue to each row so that the model knows what the actual result was so that he can learn the behaviour
  ) %>% 
  tidyr::drop_na(pred_date) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(
    product_category,
    purchase_yearweek,
    pred_date,
    future_revenue,
    total_revenue,
    lag_revenue,
    total_sellers,
    avg_reviews,
    late_proportion,
    rolling_sd4,
    rolling_mean4
  )
```

```{r}
pred_category_tbl
```

### Predictors' Correlation Matrix

```{r}
pred_category_tbl %>% 
  dplyr::select(-c(1:4)) %>% 
stats::cor(use = "complete.obs") %>% 
  corrplot::corrplot(
    method = "color",
    type = "upper",
    tl.col = "black",
    tl.srt = 45,
    order= "hclust",
    addCoef.col = "black"
  )
```

```{r}
car::vif(
  stats::lm(
    future_revenue ~ total_revenue + lag_revenue + total_sellers + avg_reviews + late_proportion + rolling_sd4 + rolling_mean4,
    data = pred_category_tbl
    )
  ) %>% 
  tibble::enframe(name = "Predictor", value = "VIF") %>% 
  gt::gt() %>% 
  gt::tab_header(title = "VIF Table") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  )
```

The correlation matrix and the VIF table above shows strong correlation between some of the predictors. Therefore, to avoid multicollinearity we are going to use PCA.

### Linear Regression Model

```{r}
set.seed(123)

dependent <- names(pred_category_tbl)[4] # This is what we're trying to predict. The y
independents <- names(pred_category_tbl)[-(1:4)] # The predictors

regression_models <- dplyr::tibble(model = "") # Created an empty table so I can add to it later


for (i in 1:length(independents)){
  betas <- paste(independents[1:i], collapse = "+")
  regression_models[i, 1] <- glue::glue("{dependent} ~ {betas}")
}


# Sampling
split_data <- pred_category_tbl %>% 
  rsample::initial_time_split(prop = 0.7)

training_data <- rsample::training(split_data)
testing_data <- rsample::testing(split_data)
```

```{r}
pca_results <- dplyr::tibble()


for (i in 1:7){
  # Creating the PCA model formula
  betas <- paste(as.character(paste0("PC", seq(1, i))), collapse = "+")
  pca_formula <- as.formula(glue::glue("{dependent} ~ {betas}"))
  
  # PCA Feature Engineering
  pca_recipe <- recipes::recipe(
    future_revenue ~ total_revenue + lag_revenue + total_sellers + avg_reviews + late_proportion + rolling_sd4 + rolling_mean4,
    data = training_data
  ) %>% 
    recipes::step_normalize(all_numeric_predictors()) %>% 
    recipes::step_pca(recipes::all_numeric_predictors(), num_comp = i) %>% 
    recipes::prep(training = training_data)
  
  training_baked_pca <- recipes::bake(pca_recipe, new_data = training_data)
  testing_baked_pca <- recipes::bake(pca_recipe, new_data = testing_data)
  
  # PCA Modeling - Linear Regression
  model_lm_pca <-
    parsnip::linear_reg(mode = "regression") %>% 
    parsnip::set_engine("lm") %>% 
    parsnip::fit(pca_formula, data = training_baked_pca)
  
  # PCA Predicting
  predicted_lm_pca <- model_lm_pca %>% 
    stats::predict(new_data = testing_baked_pca) %>% 
    dplyr::bind_cols(future_revenue = testing_baked_pca %>% dplyr::select(future_revenue))
  
  # Metrics
  glance_lm_pca <- parsnip::glance(model_lm_pca) %>% 
    dplyr::select(1:5)
  
  yardstick_lm_pca <- predicted_lm_pca %>% 
    yardstick::metrics(truth = future_revenue, estimate = .pred) %>% 
    dplyr::select(.metric, .estimate) %>% 
    tidyr::pivot_wider(
      names_from = .metric,
      values_from = .estimate
    )
  
  # Saving The Results
  pca_results <- pca_results %>% 
    dplyr::bind_rows(
      dplyr::bind_cols(
        dplyr::tibble(
          mode = glue::glue("{dependent} ~ {betas}"),
          num_comp = i
        ),
        glance_lm_pca,
        yardstick_lm_pca
      )
    )
}
```

```{r}
pca_results %>% 
  gt::gt() %>% 
  gt::tab_header(title = "PCA Linear Regression Table") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  ) %>% 
  gt::cols_align(align = "left", columns = mode)
```

```{r}
# PCA Feature Engineering
pca_recipe <- recipes::recipe(
  future_revenue ~ total_revenue + lag_revenue + total_sellers + avg_reviews + late_proportion + rolling_sd4 + rolling_mean4,
  data = training_data
) %>% 
  recipes::step_normalize(all_numeric_predictors()) %>% 
  recipes::step_pca(recipes::all_numeric_predictors(), num_comp = 5) %>% 
  recipes::prep(training = training_data)

training_baked_pca <- recipes::bake(pca_recipe, new_data = training_data)
testing_baked_pca <- recipes::bake(pca_recipe, new_data = testing_data)

# PCA Modeling - Linear Regression
model_lm_pca <-
  parsnip::linear_reg(mode = "regression") %>% 
  parsnip::set_engine("lm") %>% 
  parsnip::fit(future_revenue ~ ., data = training_baked_pca)

# PCA Predicting
predicted_lm_pca <- model_lm_pca %>% 
  stats::predict(new_data = testing_baked_pca) %>% 
  dplyr::bind_cols(
    testing_data %>% 
      dplyr::select(product_category, purchase_yearweek, pred_date, future_revenue)
  )

# Metrics
tidy_lm_pca <- parsnip::tidy(model_lm_pca) %>% 
  gt::gt()

glance_lm_pca <- parsnip::glance(model_lm_pca) %>% 
  gt::gt() %>% 
  gt::fmt_percent(columns = matches("r.squared|adj.r.squared"), decimals = 2)

yardstick_lm_pca <- predicted_lm_pca %>% 
  yardstick::metrics(truth = future_revenue, estimate = .pred) %>% 
  dplyr::arrange(.metric) %>%  
  dplyr::select(-.estimator) %>% 
  gt::gt()
```

```{r, include=FALSE}
tidy_lm_pca
glance_lm_pca
yardstick_lm_pca
```

```{r, eval=FALSE}
car::vif(model_lm_pca$fit)
```

```{r}
pca_scree <- training_data %>% 
  dplyr::select(-c(1:4)) %>% 
  stats::prcomp(
    scale = TRUE,
    center = TRUE,
    rank = 7
  ) %>% 
  factoextra::fviz_eig(
    choice = "variance",
    addlabels = TRUE,
    ncp = 7
  )

pca_scree
```

```{r}
model_lm <- regression_models %>% 
  dplyr::mutate(
    recipe_pipeline = purrr::map( # Feature Engineering
      model,
      ~ recipes::recipe(stats::as.formula(.x), data = training_data) %>% 
        recipes::step_normalize(all_numeric_predictors()) %>% 
        recipes::prep()
    ),
    training_baked = purrr::map(recipe_pipeline, ~ recipes::bake(.x, new_data = training_data)),
    testing_baked = purrr::map(recipe_pipeline, ~ recipes::bake(.x, new_data = testing_data)),
    model_lm = purrr::map2( # Modeling - Linear Regression
      model,
      training_baked,
      ~ parsnip::linear_reg(mode = "regression") %>% 
        parsnip::set_engine("lm") %>% 
        parsnip::fit(stats::as.formula(.x), data = .y)
    ),
    predicted_lm = purrr::map2( # Predicting
      model_lm,
      testing_baked,
      ~ stats::predict(.x, new_data = .y) %>% 
        dplyr::bind_cols(
          testing_data %>% 
            dplyr::select(product_category, purchase_yearweek, pred_date, future_revenue)
        )
    ),
    statistics_1 = purrr::map( # Metric 1
      model_lm, broom::glance
      )
  ) %>% 
  tidyr::unnest(statistics_1) %>% 
  dplyr::select(1:11) %>% 
  dplyr::mutate(
    statistics_2 = purrr::map( # Metric 2
      model_lm,
      parsnip::tidy
    ),
    statistics_3 = purrr::map( # Metric 3
      predicted_lm,
      ~ yardstick::metrics(.x, truth = future_revenue, estimate = .pred) %>% 
        dplyr::arrange(.metric)
    )
  ) %>% 
  dplyr::mutate_if(is.numeric, round, 3) %>% 
  tidyr::unnest(statistics_3) %>% 
  tidyr::pivot_wider(
    names_from = .metric,
    values_from = .estimate
  ) %>% 
  dplyr::select(-.estimator)
```

```{r}
model_lm
```

Based on the results above we can see that $R^2$ and $R^2_{Adjusted}$ on the training data are best on the last model with all the predictors. However, on the testing data, unseen data, the best $R^2$ is the one using the total and lag revenue as predictors. We can also see that it has a better RMSE than the last model in the training dataset.

### Regression Tree Model

Since trees split one variable at a time they can handle correlated predictors much better than linear regression.

```{r}
# PCA Modeling - Regression Tree
model_tree_pca <-
  parsnip::decision_tree(
    mode = "regression",
    tree_depth = 5,
    min_n = 5,
    cost_complexity = 0.001
    ) %>% 
  parsnip::set_engine("rpart") %>% 
  parsnip::fit(future_revenue ~ ., data = training_baked_pca)

# PCA Predicting
predicted_tree_pca <- model_tree_pca %>% 
  stats::predict(new_data = testing_baked_pca) %>% 
  dplyr::bind_cols(
    testing_data %>% 
      dplyr::select(product_category, purchase_yearweek, pred_date, future_revenue)
    )

# Metrics
yardstick_tree_pca <- predicted_tree_pca %>% 
  yardstick::metrics(truth = future_revenue, estimate = .pred) %>% 
  dplyr::arrange(.metric) %>% 
  dplyr::select(-.estimator) %>% 
  gt::gt()
```

```{r}
yardstick_tree_pca
```

```{r, fig.width= 20, fig.height=8}
rpart.plot::rpart.plot(
  model_tree_pca$fit,
  roundint = FALSE,
  cex = 1,
  fallen.leaves = FALSE,
  extra = "auto",
  main = "Regression Tree"
)
```

```{r}
tree_grid <- tidyr::expand_grid(
  tree_depth = c(3, 5, 8, 10),
  min_n = c(2, 5, 10, 20),
  cost_complexity = c(0.0001, 0.001, 0.01)
)

tree_models_grid <- tidyr::crossing(
  regression_models,
  tree_grid
)
```

```{r}
model_tree <- tree_models_grid %>% 
  dplyr::mutate(
    recipe_pipeline = purrr::map( # Feature Engineering
      model,
      ~ recipes::recipe(stats::as.formula(.x), data = training_data) %>% 
        recipes::step_normalize(all_numeric_predictors()) %>% 
        recipes::prep()
    ),
    training_baked = purrr::map(recipe_pipeline, ~ recipes::bake(.x, new_data = training_data)),
    testing_baked = purrr::map(recipe_pipeline, ~ recipes::bake(.x, new_data = testing_data)),
    model_tree = purrr::pmap( # Modeling - Regression Tree
      list(model, training_baked, tree_depth, min_n, cost_complexity),
      ~ parsnip::decision_tree(
        mode = "regression",
        tree_depth = ..3,
        min_n = ..4,
        cost_complexity = ..5
        ) %>% 
        parsnip::set_engine("rpart") %>% 
        parsnip::fit(stats::as.formula(..1), data = ..2)
    ),
    predicted_tree = purrr::map2( # Predicting
      model_tree,
      testing_baked,
      ~ stats::predict(.x, new_data = .y) %>% 
        dplyr::bind_cols(
          testing_data %>% 
            dplyr::select(product_category, purchase_yearweek, pred_date, future_revenue)
        )
    )
  ) %>%  
  dplyr::mutate(
    statistics_3 = purrr::map(
      predicted_tree,
      ~ yardstick::metrics(.x, truth = future_revenue, estimate = .pred) %>% 
        dplyr::arrange(.metric)
    )
  ) %>% 
  dplyr::mutate_if(is.numeric, round, 3) %>% 
  tidyr::unnest(statistics_3) %>% 
  tidyr::pivot_wider(
    names_from = .metric,
    values_from = .estimate
  ) %>% 
  dplyr::select(-.estimator) %>% 
  dplyr::arrange(desc(rsq))
```

```{r}
model_tree
```

Looking at the results from the decision tree we can see that the best model is the simplest model

```{r, fig.width= 20, fig.height=8}
best_tree <- model_tree %>% 
  dplyr::slice_max(rsq, n = 1)

rpart.plot::rpart.plot(
  best_tree$model_tree[[1]]$fit,
  roundint = FALSE,
  cex = 1,
  fallen.leaves = FALSE,
  extra = "auto",
  main = "Regression Tree"
)
```

```{r}
car::vif(
  stats::lm(
    future_revenue ~ total_revenue + lag_revenue,
    data = pred_category_tbl
  )
) %>% 
  tibble::enframe(name = "Predictor", value = "VIF") %>% 
  gt::gt() %>% 
  gt::tab_header(title = "VIF Table") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  )
```

### Random Forest Model

```{r}
# PCA Modeling - Random Forest
model_rf_pca <-
  parsnip::rand_forest(
    mode = "regression",
    trees = 750, # 750 different decision trees
    mtry = 3, # number of variables to randomly select and choosing the best split out of them
    min_n = 5 # Min number of observations in a leaf or end of a tree branch
  ) %>% 
  parsnip::set_engine(
    "ranger",
    importance = "impurity" # Using reduction in variance for each split
  ) %>% 
  parsnip::fit(future_revenue ~ ., data = training_baked_pca)

# PCA Predicting
predicted_rf_pca <- model_rf_pca %>% 
  stats::predict(new_data = testing_baked_pca) %>% 
  dplyr::bind_cols(
    testing_data %>% 
      dplyr::select(product_category, purchase_yearweek, pred_date, future_revenue)
  )

# Metrics
yardstick_rf_pca <- predicted_rf_pca %>% 
  yardstick::metrics(truth = future_revenue, estimate = .pred) %>% 
  dplyr::arrange(.metric) %>% 
  dplyr::select(-.estimator) %>% 
  gt::gt()
```

```{r}
yardstick_rf_pca
```

```{r}
rf_grid <- tidyr::expand_grid(
  trees = c(500, 750, 1000),
  mtry = 1:7,
  min_n = c(2, 5, 10, 15, 20)
)

rf_models_grid <- tidyr::crossing(
  regression_models,
  rf_grid
) %>% 
  dplyr::mutate(
    n_predictors = stringr::str_count(model, "\\+") + 1
  ) %>% 
  dplyr::filter(
    mtry <= n_predictors
  )
```

```{r}
model_rf <- rf_models_grid %>% 
  dplyr::mutate(
    recipe_pipeline = purrr::map( # Feature Engineering
      model,
      ~ recipes::recipe(stats::as.formula(.x), data = training_data) %>% 
        recipes::step_normalize(all_numeric_predictors()) %>% 
        recipes::prep()
    ),
    training_baked = purrr::map(recipe_pipeline, ~ recipes::bake(.x, new_data = training_data)),
    testing_baked = purrr::map(recipe_pipeline, ~ recipes::bake(.x, new_data = testing_data)),
    model_rf = purrr::pmap( # Modeling - Random Forest
      list(model, training_baked, trees, mtry, min_n),
      ~ parsnip::rand_forest(
        mode = "regression",
        trees = ..3,
        mtry = ..4,
        min_n = ..5
      ) %>% 
        parsnip::set_engine(
          "ranger",
          importance = "impurity"
        ) %>% 
        parsnip::fit(stats::as.formula(..1), data = ..2)
    ),
    predicted_rf = purrr::map2( # Predicting
      model_rf,
      testing_baked,
      ~ stats::predict(.x, new_data = .y) %>% 
        dplyr::bind_cols(
          testing_data %>% 
            dplyr::select(product_category, purchase_yearweek, pred_date, future_revenue)
        )
    )
  ) %>%  
  dplyr::mutate(
    statistics_3 = purrr::map(
      predicted_rf,
      ~ yardstick::metrics(.x, truth = future_revenue, estimate = .pred) %>% 
        dplyr::arrange(.metric)
    )
  ) %>% 
  dplyr::mutate_if(is.numeric, round, 3) %>% 
  tidyr::unnest(statistics_3) %>% 
  tidyr::pivot_wider(
    names_from = .metric,
    values_from = .estimate
  ) %>% 
  dplyr::select(-.estimator) %>% 
  dplyr::arrange(desc(rsq))
```

```{r}
model_rf
```

```{r}
model_rf %>% 
  dplyr::slice_head(n = 5) %>% 
  gt::gt() %>% 
  gt::tab_header(title = "Top 5 Random Forest Results") %>% 
  gt::tab_style(
    style = cell_fill(color = "indianred"),
    locations = cells_title(groups = "title")
  )
```
